---
title: "Boppana_Swingle_Study 1"
author: "Sriram Boppana and Connor Swingle"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    code_folding: show
---

#1 -  Setup in R
```{r setup R and load packages, message=FALSE}
knitr::opts_chunk$set(comment=NA)
options(width = 70)

## add additional libraries/packages here, as needed
## leaving the tidyverse as the last package loaded
library(readxl); library(hms); library(stringr); library(Epi); library(vcd)
library(forcats); library(Hmisc); library(gridExtra); library(pander)
library(tidyverse)
source("Love-boost.R")
```


#2 -  What is this? 

#3 - Importing the datasets
```{r importing data}
surv_raw <- read_excel("431-2017-project-survey-responses-raw-deid.xlsx")

names_sur <- read_excel("431-2017-project-survey-names.xlsx")

surv_tidy <- surv_raw

names(surv_tidy) <- names_sur$short_name

```

```{r build_scales_etc}
surv_tidy$Q051_Gould_score <- round(rowSums(surv_tidy[,52:60])/9,3)
```

#4 - Analysis 1 - Comparing Means of Two Populations

##4.1 Boxplot Analysis
```{r Analysis 1.1 Boxplot}
by(surv_tidy$caffeine_Q021, surv_tidy$married_Q012, Hmisc::describe)

married <- surv_tidy %>%
  filter(married_Q012 == "Yes")
unmarried <- surv_tidy %>%
  filter(married_Q012 == "No")


ggplot(surv_tidy, aes(x = surv_tidy$married_Q012, y = surv_tidy$caffeine_Q021, fill = surv_tidy$married_Q012)) +
    geom_boxplot() +
    geom_jitter(width = 0.1, height = 0, color = "orangered") +
    guides(fill = FALSE) + 
    labs(title = "Boxplot of Cups of Caffinated Beverages ingested by Married and Unmarried Individuals in One Week",
         x = "Married?", y = "Cups of Caffinated Beverages") + 
    coord_flip() +
    theme_bw()

```
##4.2 Histogram Analysis
```{r Analysis 1.2 Histogram}
ggplot(surv_tidy, aes(x = caffeine_Q021, fill = married_Q012)) +
    geom_histogram(color = "white", bins = 20) +
    guides(fill = FALSE) + 
    labs(title = "Histograms of Cafeinated Beverages taken by married and Unmarried Individuals",
         x = "Cups of Cafeinated Beverages") +
    theme_bw() +
    facet_wrap(~ married_Q012)
```
```{r Analysis 1.3}
ggplot(surv_tidy, aes(x = sqrt(caffeine_Q021), fill = married_Q012)) +
    geom_histogram(color = "white", bins = 20) +
    guides(fill = FALSE) + 
    labs(title = "Histograms of Cafeinated Beverages taken by married and Unmarried Individuals",
         x = "Square root of Cups of Cafeinated Beverages") +
    theme_bw() +
    facet_wrap(~ married_Q012)
```

```{r Analysis 1.4}
ggplot(surv_tidy, aes(x = sqrt(caffeine_Q021), y = married_Q012, fill = caffeine_Q021)) +
    ggridges::geom_density_ridges(scale = 1) +
    guides(fill = FALSE) + 
    labs(title = "Caffinated Beverages taken by Married and Unmarried Individuals",
         x = "Caffeinated Beverages", y = "Married?") +
    ggridges::theme_ridges()
```

```{r Analysis 1.5}
surv_tidy %>% group_by(married_Q012) %>% 
    summarise(n = n(), 
              median = median(caffeine_Q021), 
              Q1 = quantile(caffeine_Q021, 0.25), 
              Q3 = quantile(caffeine_Q021, 0.75),
              mean = mean(caffeine_Q021),
              sd = sd(caffeine_Q021)) 
```

```{r Analysis 1.6}
bootdif(surv_tidy$caffeine_Q021, surv_tidy$married_Q012, conf.level = 0.90)
```

#5 - Analysis 2 - Comparing Means of Three or More Groups

```{r Analysis 2.1}
#Hmisc::describe(surv_tidy$birth_year_Q001)

birthyear <- surv_tidy$birth_year_Q001
ggplot(surv_tidy, aes(x = birthyear, fill = birthyear)) +
    geom_histogram(color = "white", bins = 50) +
    guides(fill = FALSE) + 
    labs(title = "Age") +
    theme_bw() +
    facet_wrap(~ married_Q012)
```

```{r Analysis 2.2}
surv_tidy$birthgroup[surv_tidy$birth_year_Q001 < 1980] <- "Before 1980"
surv_tidy$birthgroup[surv_tidy$birth_year_Q001 >= 1980 & surv_tidy$birth_year_Q001 < 1990] <- "1980-1990"
surv_tidy$birthgroup[surv_tidy$birth_year_Q001 >= 1990 & surv_tidy$birth_year_Q001 < 2000] <- "After 1990"

eda.ksam(outcome = surv_tidy$caffeine_Q021, group = surv_tidy$birthgroup, axis.title = "Caffeine by Birth group", main.title = "Birth Group vs Caffeine Intake", notch = FALSE)

```
```{r Analysis 2.3}
anova(lm(surv_tidy$caffeine_Q021 ~ surv_tidy$birthgroup))

```

#6 - Analysis 3 - Regression Comparison of Means with Adjustments

##6.1 - The Regression Model, Adjusting for a Single Quantitative Covariate
```{r Analysis 3.1}
Hmisc::describe(surv_tidy$facebook_Q035)

```

```{r Analysis 3.2}
summary(lm(caffeine_Q021 ~ birthgroup + facebook_Q035, data = surv_tidy))

```


```{r Analysis 3.3}
anova(lm(caffeine_Q021 ~ birthgroup + facebook_Q035, data = surv_tidy))

```
##6.2 - Predicting the outcome at the average level of the covariate for each group
```{r Analysis 3.4}
mean(surv_tidy$facebook_Q035)

model1 <- lm(caffeine_Q021 ~ birthgroup + facebook_Q035, data = surv_tidy)
model1
new1 <- data.frame(birthgroup = c("Before 1980", "1980-1990", "After 1990"), facebook_Q035 = rep(mean(surv_tidy$facebook_Q035), 3))
new1 # sanity check
pander(predict(model1, newdata = new1, interval = "prediction", level = 0.90))
```
```{r Analysis 3.5}
par(mfrow=c(1,2))
plot(model1, which = 1:2)
```

# 7 - Analysis 4: Two-Way (2 x 2) Contingency Table

## 7.1 - Building the 2x2 Table from data

```{r Analysis 4.1}
t1 <- table(surv_tidy$instrument_Q046, surv_tidy$smoked100_Q013)
rownames(t1) <- c("Did not play instrument","Played instrument")
colnames(t1) <- c("< 100 cigarettes", "> 100 cigarettes")
emphasize.italics.rows(3)
emphasize.italics.cols(3)
panderOptions('table.emphasize.rownames', FALSE)
pander(addmargins(t1), caption = "Musicial Instrument vs. Cigarette Smoking to 431")

```
```{r Analysis 4.2}
twoby2(t1 + 2, conf.level = 0.90) # uses Bayesian augmentation, 90% confidence level
```
# 8 - Analysis 5: Two-Way (3 x 4) Contingency Table

## 8.1 - Viewing the table
```{r Analysis 5.1}
t2 <- table(surv_tidy$computer_OS_Q071, surv_tidy$phone_OS_Q072)
rownames(t2) <- c("iOS","Windows","Chrome")
colnames(t2) <- c("iOS","Windows","Android","No smart phone")
emphasize.italics.rows(3)
emphasize.italics.cols(3)
panderOptions('table.emphasize.rownames', FALSE)
pander(addmargins(t2), caption = "Computer vs. Smart Phone Operating Systems")
```

```{r Analysis 5.2}
chisq.test(t2)
fisher.test(t2)
assoc(t2)

```

# 9 - Analysis 6: Three-Way Contingency Table

##9.1 Generating table
```{r Analysis 6.1}
t3.flat <- table(surv_tidy$gould_Q051a, surv_tidy$instrument_Q046, surv_tidy$smoked100_Q013)
dimnames(t3.flat)[[1]] <- c("Low", "LowMedium", "Medium", "MediumHigh") # specify the names in t5's first variable
dimnames(t3.flat)[[2]] <- c("Played Instrument", "No Instrument") # specify names in t5's second variable
dimnames(t3.flat)[[3]] <- c("Smoked > 100", "Smoked<100") # specify names in t5's third variable
ftable(t3.flat)

```


##9.2 Wolf Test
```{r Analysis 6.2}

woolf_test(t3.flat)
```
```{r Analysis 6.3}
mantelhaen.test(t3, conf.level = .90)
```

